Merging datasets through deep learning

### 2018-08-20 paper2018/introduction.tex

In this paper, we propose a novel data driven approach to the problem of joining different surface representations of the same entity.  Our approach relies on building supervised deep learning models to automatically learn the correct set of transformations needed to compute the equality of two cell values.  Because a deep learning model can learn the features necessary for matching from the data, it should in theory be able to generalize better to new datasets than the data driven approaches outlined in \cite{He:2015:SJS:2824032.2824036}.   Conceptually, our approach is similar in spirit to the approach described in \cite{auto-join-joining-tables-leveraging-transformations}, but we use deep neural networks to learn the right function for the join operation instead of synthesizing different programs for different entity types.  The advantage of building such a function compared to work described in \cite{auto-join-joining-tables-leveraging-transformations} is that (a) we can use this function in novel ways to completely eliminate the blocking step that is typically needed to identify the set of potentially joinable rows, as we describe below, and (b) we can use this function on arbitrary surface forms such as faces, where valid transformations of a face cannot be readily described in terms of a program.

@@

\begin{table}[!htb]
    \caption{Example of a merge problem}
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table A}
        \begin{tabular}{|l|l|}
          \hline
           Dept & Name \\
           \hline
           1    & Douglas Adams \\
           2    & John Adams \\
           3  & Douglas Bard \\
           \hline
        \end{tabular}
    \end{subtable}%
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table B}
        \begin{tabular}{|l|l|}
          \hline
           Name & Salary \\
           \hline
           Douglas N. Adams & 2000 \\
           Adams, John & 3000 \\
           John Smith & 3000 \\
           \hline
        \end{tabular}
    \end{subtable}
    \label{table-example}
\end{table}

>>>>

In this paper, we propose a novel approach to the problem of joining different surface representations of the same entity, inspired in part by recent advances in deep learning.  Our approach relies on building supervised deep learning models to automatically learn the correct set of transformations needed to compute the equality of two cell values.  Because a deep learning model can learn the features necessary for matching from the data, it should in theory be able to generalize better to new datasets than the data driven approaches outlined in \cite{He:2015:SJS:2824032.2824036}.   Conceptually, our approach is similar in spirit to the approach described in \cite{auto-join-joining-tables-leveraging-transformations}, but we use deep neural networks to learn the right function for the join operation instead of synthesizing different programs for different entity types.  The advantage of building such a function compared to work described in \cite{auto-join-joining-tables-leveraging-transformations} is that (a) we can use this function in novel ways to completely eliminate the blocking step that is typically needed to identify the set of potentially joinable rows, as we describe below, and (b) we can use this function on arbitrary surface forms such as faces, where valid transformations of a face cannot be readily described in terms of a program.

@@

\begin{table}[!htb]
    \caption{Example of a merge problem}
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table A}
        \begin{tabular}{|l|l|}
          \hline
           Dept & Name \\
           \hline
           1    & Douglas Adams \\
           2    & John Adams \\
           3  & Douglas Bard \\
           \hline
        \end{tabular}
    \end{subtable}%
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table B}
        \begin{tabular}{|l|l|}
          \hline
           Name & Salary \\
           \hline
           Douglas Noel Adams & 2000 \\
           Adams, John & 3000 \\
           John Smith & 3000 \\
           \hline
        \end{tabular}
    \end{subtable}
    \label{table-example}
\end{table}

### 2018-08-20 paper2018/joins.tex

\section{Applying deep learning models to joins}
\label{join_system}
\begin{figure}
\includegraphics[width=1.0\linewidth]{joins}
\caption{Applying deep learning models to joins}
\label{join_fig}
\end{figure}

Figure~\ref{join_fig} provides an overview of how deep learning models can be used for merging datasets.  Panel (a) in the figure are the two datasets to be joined.  The datasets are labeled arbitrarily as a left table and a right table.  For each cell value in the two columns to be joined, we obtain vector embeddings from the siamese network that was used to estimate distance for same and different surface forms for the same entity.  Note that although the siamese network has three separate networks, each network within a triplet network is in fact identical to the other two networks because they share weights.  A single version of the network with learnt weights is used to generate the embeddings for the cell value as shown in panel (b). For the left table cell values, vector embeddings are inserted into an approximate nearest neighbors index, as shown in panel (c).  For each cell value in the right table, vector embeddings are used as `query vectors' to query the approximate nearest neighbors index as shown in panel (d).  In our context, merging the datasets would involve joining the top $k$ rows in the left table that are `closest' in distance to each cell value in the right table.  Note that the choice of $k$ clearly has a direct effect on the tradeoff between precision and recall, but this is true for any type of join algorithm that is not based on equality.

>>>>

\section{Applying deep learning models to joins}
\label{join_system}

\begin{figure*}[htb]
    \centering
    \begin{subfigure}[t]{0.24\linewidth}
        \centering
        \includegraphics[width=.9\linewidth]{join1}
        \caption{Columns to be merged}
        \label{join1}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.24\linewidth}
        \centering 
        \includegraphics[width=.9\linewidth]{join2}
        \caption{Create embeddings for each cell value in each column using the siamese model}
        \label{join2}
    \end{subfigure}
    ~ 
    \begin{subfigure}[t]{0.24\linewidth}
        \centering 
        \includegraphics[width=.9\linewidth]{join3}
        \caption{Index embeddings for left table's cell values in a nearest neighbors index}
        \label{join3}
    \end{subfigure}
    ~ 
    \begin{subfigure}[t]{0.24\linewidth}
        \centering 
        \includegraphics[width=.9\linewidth]{join4}
        \caption{For each cell in right table, find closest neighbor(s) and merge these rows}
        \label{join4}
    \end{subfigure}
    \caption{Overview of joins using deep learning}
    \label{join_fig}
\end{figure*}

Figure~\ref{join_fig} provides an overview of how deep learning models can be used for merging datasets.  Figure~\ref{join1} in the figure shows the two columns in the two datasets to be joined.  The datasets are labeled arbitrarily as a left table and a right table.  For each cell value in the two columns to be joined, we obtain vector embeddings from the siamese network that was used to estimate distance for same and different surface forms for the same entity.  Note that although the siamese network has three separate networks, each network within a triplet network is in fact identical to the other two networks because they share weights.  A single version of the network with learnt weights is used to generate the embeddings for the cell value as shown in figure ~\ref{join2}. For the left table cell values, vector embeddings are inserted into an approximate nearest neighbors index, as shown in figure ~\ref{join3}.  For each cell value in the right table, vector embeddings are used as `query vectors' to query the approximate nearest neighbors index as shown in figure ~\ref{join4}.  In our context, merging the datasets would involve joining the top $k$ rows in the left table that are `closest' in distance to each cell value in the right table.  Note that the choice of $k$ clearly has a direct effect on the tradeoff between precision and recall, but this is true for any type of join algorithm that is not based on equality.

### 2018-08-20 paper2018/main.tex

\maketitle
\begin{abstract}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions. 
\end{abstract}

\input{introduction}
\input{siamese_networks}

>>>>

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging is a join between two columns that frequently have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity to allow so-called `fuzzy' joins.  We propose a novel approach to performing merges inspired in part by recent advances in deep learning.  Our approach depends on (a) creating a function that maps surface forms into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a space partitioning algorithm to find the forms that can be potentially joined together.  Our approach for merging datasets is efficient $\mathcal{O}(n\log{}n)$ and flexible because it can be applied to any data provided one can create a mapping function such that joinable surface forms of the entity are closer in vector space than disparate forms.  We demonstrate the generality of this approach by creating such a mapping function using deep learning models for people's names and company names drawn from Wikidata, and show that such a function does in fact produce vector embeddings that can be used together with space partitioning algorithms to perfom efficient joins.
\end{abstract}

\input{introduction2}
\input{siamese_networks}
\input{joins}
\input{datasets}

### 2018-08-20 paper2018/triplet_selection.tex

Approximate nearest neighbor indexes are highly efficient methods for selecting the top-K neighbors of a given vector by Euclidean distance, cosine similarity or other distance metrics.  They are based on space partitioning algorithms, such as \textit{k-d trees}, where the vector space is iteratively bisected into two disjoint regions.  Finding the nearest neighbors of a vector involves traversal of the tree to a leaf by evaluating the queried vector at each split point in the tree.  Depending on the position of the vector within the space bounded by a split point, it may need to traverse both points at a split node.  However, large portions of the tree (and hence most vectors in the dataset) never get compared to the queried vector.  The average complexity to query the vectors of a neighbor is $O(log N)$ where N is the number of vectors in the dataset.  Implementations exist now for fast, memory-efficient ANN indexes that scale up to a billion vectors \cite{JDH17} using techniques to compress vectors in memory efficiently.  In our work, we used the Annoy ANN implementation \cite{annoy_git} in our work which is based on the refinements to \textit{k-d trees} ideas described in \cite{annoy-paper}.

>>>>

Approximate nearest neighbor indexes are highly efficient methods for selecting the top-K neighbors of a given vector by Euclidean distance, cosine similarity or other distance metrics.  They are based on space partitioning algorithms, such as \textit{k-d trees}, where the vector space is iteratively bisected into two disjoint regions.  Finding the nearest neighbors of a vector involves traversal of the tree to a leaf by evaluating the queried vector at each split point in the tree.  Depending on the position of the vector within the space bounded by a split point, it may need to traverse both points at a split node.  However, large portions of the tree (and hence most vectors in the dataset) never get compared to the queried vector.  The average complexity to query the vectors of a neighbor is $O(log N)$ where N is the number of vectors in the dataset.  Implementations exist now for fast, memory-efficient ANN indexes that scale up to a billion vectors \cite{JDH17} using techniques to compress vectors in memory efficiently.  In our work, we used the Annoy ANN implementation \cite{annoy_git} in our work which is based on the refinements to \textit{k-d trees} ideas described in \cite{ann_paper}.
